---
title: "Acumen"
author: "Aung Myo Htut"
date: "2024-02-01"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#Load the libraries
library(tidyverse)
library(readxl)
library(table1)
library(naniar)
library(ggplot2)
library(openxlsx)
library(knitr)
library(ggpubr)
```

#load the file
```{r loading}
setwd("C:/Users/draun/Desktop/Acumen Exercise")
data <- read_xlsx("Acumen_Data_Analysis_Exercise_Htut_AungMyo.xlsx", sheet = 2)
data <- data[,1:9]
dim(data)
```

#Rename variables
```{r renaming}
names(data) <- c("obs_no","qr", "eid", "sex", "race","age","hosp_visit","salary","health_score")
colnames(data)
# To know internal structure of dataframe
str(data)
```

Question 1: Understanding the data.
a) Are all the values in the data reasonable? Are there missing values?
b) What are the characteristics of employees at Company A? Do these demographics change over time?
Use tables and charts to understand the data and demographic characteristics of employees at Company A. 

#Summary Statistics
```{r Summary}
summary(data)
```

#Missing values
```{r Missing Values}
#Total Missing Values
sum(is.na(data))

#Columns with missing values
colSums(is.na(data))

#percentage of missing values in the dataframe
pct_miss(data)

#Percent of rows with any missing values
pct_miss_case(data)

#Percent of rows that are complete
pct_complete_case(data)

#Visualizing missing values
gg_miss_var(data, show_pct = TRUE)
```

#Factor Variables
```{r Factoring}
data$sex <- as.factor(data$sex)
data$race <- as.factor(data$race)
data$hosp_visit <- as.factor(data$hosp_visit)

str(data)
summary(data)

```

#Removing Missing cases
```{r NAs Omit}
#Dropping NAs
comp_data <- data %>% drop_na()

#Recheck missing values
sum(is.na(comp_data))

#dimension of data
dim(comp_data)
```

#Finding and removing outliers in age
```{r age outliers}
# Under 18 
under18 <- comp_data %>% filter(age<18) %>% (n=nrow)
under18

#Over 65
over65 <- comp_data %>% filter (age>65) %>% (n=nrow)
over65

#Removing those with under 18 and over 65
comp_data <- comp_data %>% filter(age >= 18 & age <= 65)
dim(comp_data)

```

#Finding and removing outliers in health score
```{r health score outliers}
#6 is the highest health score supposed to be in the dataset
#Over 6 healthscore
sum(comp_data$health_score >6)

comp_data <- comp_data %>% filter (health_score <=6)
dim(comp_data)
```


#Table - Overall
```{r table overall}
#Overall
t1 <- table1(~ sex + race + age + hosp_visit + salary + health_score,
             data =  comp_data,
             overall = "Total")
t1
#write.xlsx(t1,"table1.xlsx")
```


#Histograms
```{r age, salary and health score}
age_his <- ggplot(comp_data, aes(x = age)) +
  geom_histogram(binwidth = 1, color = "black", fill = "white")+
  geom_vline(aes(xintercept = mean(age)), color = "red", linewidth = 0.8)+
  geom_vline(aes(xintercept = median(age)), color = "blue", linewidth = 0.8)+
  xlab("Age of Employees at Company A") +
  ylab("Count") +
  ggtitle ("Histogram of Age of Employees at Company A")
age_his
#ggsave(filename = "age_his.png", plot = age_his,width = 5, height = 4, units = "in",bg = "white")

salary_his <- ggplot(comp_data, aes(x = salary)) +
  geom_histogram(bins= 30, color = "black", fill = "white")+
  geom_vline(aes(xintercept = mean(salary)), color = "red", linewidth = 0.8)+
  geom_vline(aes(xintercept = median(salary)), color = "blue", linewidth = 0.8, linetype = "dashed")+
  xlab("Salary of Employees at Company A") +
  ylab("Count") +
  ggtitle ("Histogram of Salary of Employees at Company A")
salary_his
#ggsave(filename = "salary_his.png", plot = salary_his,width = 5, height = 4, units = "in",bg = "white")


healthscore_his <- ggplot(comp_data, aes(x = health_score)) +
  geom_histogram(bins= 35, color = "black", fill = "white")+
  geom_vline(aes(xintercept = mean(health_score)), color = "red", linewidth = 0.8)+
  geom_vline(aes(xintercept = median(health_score)), color = "blue", linewidth = 0.8, linetype = "dashed")+
  xlab("Health Score of Employees at Company A") +
  ylab("Count") +
  ggtitle ("Histogram of Health Score of Employees at Company A")
healthscore_his
#ggsave(filename = "healthscore_his.png", plot = healthscore_his,width = 5, height = 4, units = "in",bg = "white")
```


#Demographics change over time
```{r demo change over time}
demo_overtime <- comp_data %>%
  group_by(qr) %>%
  summarise(mean_age = mean(age),
            mean_salary = mean(salary),
            hospvisit_sum = sum(hosp_visit=="1"),
            mean_hs = mean(health_score))
demo_overtime

```


#Create visualization of Age over time
```{r Age Over Time}
aot <- ggplot(data = demo_overtime, aes(x = qr, y = mean_age)) +
  geom_line(color = "red") +
  scale_x_continuous(breaks = c(seq(1,12,1))) +
  labs(title = "Mean Age of Employees Over Time at Company A",
       y = "Mean Age",
       x = "Quarter") +
  theme_bw() +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 12))
aot
#ggsave(filename = "aot.png", plot = aot,width = 5, height = 4, units = "in",bg = "white")
```

#Create visualization of Salary over time
```{r Salary Over Time}
sot <- ggplot(data = demo_overtime, aes(x = qr, y = mean_salary)) +
  geom_line(color = "red") +
  scale_x_continuous(breaks = c(seq(1,12,1))) +
  labs(title = "Mean Salary of Employees Over Time at Company A",
       y = "Mean Salary",
       x = "Quarter") +
  theme_bw() +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 12))
sot
#ggsave(filename = "sot.png", plot = sot,width = 5, height = 4, units = "in",bg = "white")
```

#Create visualization of Hospital Visits over time
```{r Hospital Visits Over Time}
hvot <- ggplot(data = demo_overtime, aes(x = qr, y = hospvisit_sum)) +
  geom_col(fill = "steelblue") +
  geom_text(aes(label = hospvisit_sum), vjust = -0.5, color = "red", size = 3) +
  scale_x_continuous(breaks = c(seq(1,12,1))) +
  labs(title = "Hospital Visits of Employees Over Time",
       y = "Hospital Visits",
       x = "Quarter") +
  theme_bw() +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 12))
hvot
#ggsave(filename = "hvot.png", plot = hvot,width = 5, height = 4, units = "in",bg = "white")
```

#Health Score Trend Over Time
```{r health score over time}
hsot <- ggplot(data = demo_overtime, aes(x = qr, y = mean_hs)) +
  geom_line(color = "red") +
  scale_x_continuous(breaks = c(seq(1,12,1))) +
  labs(title = "Mean Health Score of Employees Over Time at Company A",
       y = "Mean Health Score",
       x = "Quarter") +
  theme_bw() +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 12))
hsot
#ggsave(filename = "hsot.png", plot = hsot,width = 5, height = 4, units = "in",bg = "white")

```



Question 2: Exploring Relationships
a) Which characteristics are associated with the health score? 
Use tables and charts (suggestion: scatter plots for continuous variables) to determine which characteristics are associated with the health score.

#Normality and Homogeneity Assumption
```{r}
library(nortest)
ad.test(comp_data$health_score)
ad.test(comp_data$age)
ad.test(comp_data$salary)
library(ggpubr)
ggqqplot(comp_data$health_score, main = "Q-Q Plot for Health Score with 95% Confidence Interval")
ggqqplot(comp_data$age, main = "Q-Q Plot for Age with 95% Confidence Interval")
ggqqplot(comp_data$salary, main = "Q-Q Plot for Salary with 95% Confidence Interval")
library(car)
leveneTest(health_score ~ sex, data =  comp_data)
leveneTest(health_score ~ race, data = comp_data)
```


#Sex and Health Score
```{r sex and health score}
#Box plot
sexvshs <- ggplot(comp_data, aes(x = sex, y = health_score, fill = sex)) + 
  geom_boxplot() +
  scale_y_continuous(breaks = c(seq(1,6,1)))+
  labs(title = "Sex vs Health Score",
       x = "Sex",
       y = "Health Score")
sexvshs
#t.test(health_score ~ sex, data= comp_data)
wilcox.test(health_score ~ sex, data = comp_data)

#ggsave(filename = "sexvshs.png", plot = sexvshs,width = 5, height = 4, units = "in",bg = "white")

# t2 <- table1 (~ health_score | sex, data = comp_data, overall = "Total")
# t2
# file_path <- "C:/Users/draun/Desktop/Acumen Exercise/table1.xlsx"
# wb <- loadWorkbook(file_path)
# writeData(wb, sheet = "Sheet2",t2)
# saveWorkbook(wb,file_path,overwrite = TRUE)
```

#Race and Health Score
```{r Race vs Health Score}
racevshs <- ggplot(comp_data, aes(x = race, y = health_score, fill = race)) + 
  geom_boxplot() +
  scale_y_continuous(breaks = c(seq(1,6,1))) +
  labs(title = "Race vs Health Score",
       x = "Race",
       y = "Health Score")
racevshs
result <- aov(health_score ~ race, data = comp_data)
summary(result)
kruskal.test(health_score ~ race, data = comp_data)

#ggsave(filename = "racevshs.png", plot = racevshs,width = 5, height = 4, units = "in",bg = "white")

# t3 <- table1 (~ health_score | race, data = comp_data, overall = "Total")
# t3
# file_path <- "C:/Users/draun/Desktop/Acumen Exercise/table1.xlsx"
# wb <- loadWorkbook(file_path)
# writeData(wb, sheet = "Sheet3",t3)
# saveWorkbook(wb,file_path,overwrite = TRUE)
```


# Age and Health Score
```{r age vs health score}
agevshs <- ggplot(comp_data, aes(x = age, y = health_score)) +
  geom_point(color = "blue", size = 1.5, alpha = 0.3) +
  geom_smooth(method = lm, se = FALSE, color = "red") +
  scale_x_continuous(breaks = c(seq(15,70,5))) +
  scale_y_continuous(breaks = c(seq(1,6,1))) +
  labs(title = "Age vs Health Score",
       x = "Age",
       y = "Health Score") 
agevshs
cor.test(comp_data$age, comp_data$health_score, method = "spearman")
#ggsave(filename = "agevshs.png", plot = agevshs,width = 5, height = 4, units = "in",bg = "white")
  
```

#Hosptal Visit vs Health Score
```{r Hosp Visit vs Health Score}
hvvshs <- ggplot(comp_data, aes(x = hosp_visit, y = health_score, fill = hosp_visit)) + 
  geom_boxplot() +
  scale_y_continuous(breaks = c(seq(1,6,1)))+
  labs(title = "Hospital Visit vs Health Score",
       x = "Hospital Visit",
       y = "Health Score")
hvvshs
wilcox.test(health_score ~ hosp_visit, data = comp_data)
#ggsave(filename = "hvvshs.png", plot = hvvshs,width = 5, height = 4, units = "in",bg = "white")

# t4 <- table1 (~ health_score | hosp_visit, data = comp_data, overall = "Total")
# t4
# file_path <- "C:/Users/draun/Desktop/Acumen Exercise/table1.xlsx"
# wb <- loadWorkbook(file_path)
# writeData(wb, sheet = "Sheet4",t4)
# saveWorkbook(wb,file_path,overwrite = TRUE)
```

#Salary vs Health Score
```{r Salary vs Health Score}
salaryvshs <- ggplot(comp_data, aes(x = salary, y = health_score)) +
  geom_point(color = "blue", size = 1.5, alpha = 0.3) +
  geom_smooth(method = lm, se = FALSE, color = "red") +
  scale_y_continuous(breaks = c(seq(1,6,1))) +
  labs(title = "Salary vs Health Score",
       x = "Salary",
       y = "Health Score") 
salaryvshs
#ggsave(filename = "salaryvshs.png", plot = salaryvshs,width = 5, height = 4, units = "in",bg = "white")

cor.test(comp_data$salary, comp_data$health_score)
```



